---
title: "Week2HW"
output: html_document
date: "2022-09-07"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 0

-   Draw a concepts diagram that uses all the following Github
    terminology
    -   Include any missing keywords that will simplify the concepts
        diagrams

`Push`, `Repository`, `Clone`, `Pull`, `Pull Request`, `Branch`,
`Merging`, `Github Client`, `README file`, `Private or Public`

``` {.R}
# Add your photo here
```

#### Question 1 {#question-1}

`matrix(c(1,2,3,4,5,6) , nrow = 3)`

-   Running the expression produces the following matrix

  --- ---
  1   4
  2   5
  3   6
  --- ---

-   How can you modify the call to `matrix()` to produce the following
    matrix instead?

          col_1   col_2
  ------- ------- -------
  row_1   1       2
  row_2   3       4
  row_3   5       6

-   Note that you need to name the columns (col_1 and col_2) and name
    the rows (row_1, row_2, row_3)

Hint: Use the `?` symbol to invoke the matrix documentation

```{r}
matrix_1 <- matrix(c(1,2,3,4,5,6) , nrow = 3)
rownames(matrix_1) <- c("row_1", "row_2", "row_3")
colnames(matrix_1) <- c("col_1", "col_2")
matrix_1

```

#### Question 2

-   Load then sort the airquality data frame on its `Temp` and `Solar.R`
    columns in reverse order (largest to smallest values)
    -   The function to sort a data frame is called order
-   Display only the first 15 lines of your table

```{r}
data("airquality")
AQ <- airquality

subsetAQ <- AQ[,c( "Temp","Solar.R")]

Temp_sorted <- subsetAQ[order(subsetAQ$Temp, decreasing = TRUE),]
head(Temp_sorted, 15)


Solar.R_sorted <- subsetAQ[order(subsetAQ$Solar.R, decreasing=TRUE),]
head(Solar.R_sorted, 15)


subsetAQ_sorted <- subsetAQ[order(subsetAQ$Temp,subsetAQ$Solar.R,decreasing = TRUE),]
head(subsetAQ_sorted, 15)
```


#### Question 3

-   Sort the airquality data frame on its `Temp` in decreasing order and
    `Solar.R` in increasing order
-   Display only the first 15 lines of your table

```{r}
#Temp_sorted2 <- subsetAQ[order(subsetAQ$Temp, decreasing = TRUE),]
#head(Temp_sorted2,15)

#Solar.R_sorted2<- subsetAQ[order(subsetAQ$Solar.R, decreasing = FALSE),]
#head(Solar.R_sorted2, 15)

#I had to use tidyverse, I don't know how to solve this question with regular commands
library("tidyverse")

subsetAQ2 <- subsetAQ %>% 
select(Temp, Solar.R) %>% 
distinct(Temp, Solar.R) %>% 
arrange(., Solar.R) %>% 
arrange(., desc(Temp)) 

head(subsetAQ2, 15)
```

#### Question 4

-   There are various ways to select a subset of observations from a
    data frame.

-   Consult your [R Reference
    Card](https://cran.r-project.org/doc/contrib/Baggott-refcard-v2.pdf),
    see `Data Selection and Manipulation` section.

    -   What operations can you use to select all observations where the
        temperature is 72. Give at least two different answers to this
        question
        
        
```{r}
head(AQ)

AQ[AQ$Temp == 72,] #we can use logical operators to find the value that we want with a direct equality


AQ[AQ$Temp < 73 & AQ$Temp >71,] #we can also use intervals to identify a range of interest 
```

#### Question 6

-   You may have noticed when working with the `airqulity` data that
    some values show as `NA`
-   `NA` stands for not available, or missing values.
-   A major part of data wrangling consists of cleaning missing values
    by either:
    -   Dropping the lines that have missing values
    -   Sometimes we can drop the column with missing values if the
        column is made of predominantly missing values
    -   Imputing the missing values, which uses educated guesses (or
        more complex algorithms) to fill the missing values
-   Find and remove all rows that are missing values for the `Solar.R`
    or `Ozone` variables
-   Save the cleaned data to a new data frame called airquality_no_na
    -   How many lines have been removed?
    
    
```{r}
head(AQ[is.na(AQ$Ozone),]) #we want to identify the places in the dataset that have NA values

airquality_no_na <- AQ[!is.na(AQ$Ozone),] #we remove the NA values from the Ozone column in the dataset
head(airquality_no_na)

nrow(AQ) 
nrow(airquality_no_na)
nrow(AQ) - nrow(airquality_no_na) #we removed 37 rows
```

#### Question 7

-   Let\'s use a different strategy and impute the missing value.
    -   replace the missing values for Solar.R using that month\'s
        average.
    -   Example:
        -   The missing value for line 6 should be replaced with the
            average for month 5.
        -   The missing value for line 97 should be replaced with the
            average for month 8.
            
```{r}
AQ[is.na(AQ$Solar.R),] #we identify the places where we need to replace the NA values


#Replacing the values for May 

May_SR <- AQ[AQ$Month == 5 &!is.na(AQ$Solar.R),] 
#we isolate the values in the dataset
#we are focusing on the month of May and 
#we remove the NA values from the Solar.R column so we can later take the average

May_avgSR <- mean(May_SR$Solar.R) #we take the average for May

AQ[is.na(AQ$Solar.R) & AQ$Month == 5,] <- May_avgSR 
#we replace the NA values in May by the average Solar Radiation of May


#Replacing the values for August 

Aug_SR <- AQ[AQ$Month == 8 &!is.na(AQ$Solar.R),] 
#isolate the values of Solar.R in august and remove the NA so we can take the average

Aug_avgSR <- mean(Aug_SR$Solar.R) 
Aug_avgSR
#we take the average Solar.R in august

AQ[is.na(AQ$Solar.R) & AQ$Month == 8,] <- Aug_avgSR
#we replace the NA values in august with the average Solar.R in august

AQ[is.na(AQ$Solar.R), ] #we check for the remaining NA 
# we notice that there is no more NA values to replace

head(AQ, 15)
```




